Response to reviews

### 2023-02-21 manuscript/manuscript.tex
provide a better abstract

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[english,man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={DateLife: leveraging databases and analytical tools to reveal the dated Tree of Life},
            pdfauthor={Luna L. Sánchez Reyes1,2, Emily Jane McTavish1, \& Brian O'Meara2},
            pdfkeywords={Tree; Phylogeny; Scaling; Dating; Ages; Divergence times; Open Science; Congruification; Supertree; Calibrations; Secondary calibrations},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{Tree; Phylogeny; Scaling; Dating; Ages; Divergence times; Open Science; Congruification; Supertree; Calibrations; Secondary calibrations\newline\indent Word count: 5629}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

@@

\abstract{%
Achieving a high-quality reconstruction of a phylogenetic tree with branch lengths proportional to absolute time (chronogram) is a difficult and time-consuming task. But the increased availability of fossil and molecular data, and time-efficient analytical techniques has resulted in many recent publications of large chronograms for a large number and wide diversity of organisms.
Knowledge of the evolutionary time frame of organisms is key for research in the natural sciences. It also represent valuable information for education, science communication, and policy decisions.
When chronograms are shared in public and open databases, this wealth of expertly-curated and peer-reviewed data on evolutionary timeframe is exposed in a programatic and reusable way, as intensive and localized efforts have improved data sharing practices, as well as incentivizited open science in biology.
Here we present DateLife, a service implemented as an R package and an R Shiny website application available at www.datelife.org, that provides functionalities for efficient and easy finding, summary, reuse, and reanalysis of expert, peer-reviewed, public data on time frame of evolution.
The main DateLife workflow constructs a chronogram for any given combination of taxon names by searching a local chronogram database constructed and curated from the Open Tree of Life Phylesystem phylogenetic database, which incorporates phylogenetic data from the TreeBASE database as well.
We implement and test methods for summarizing time data from multiple source chronograms using supertree and congruification algorithms, and using age data extracted from source chronograms as secondary calibration points to add branch lengths proportional to absolute time to a tree topology.
DateLife will be useful to increase awareness of the existing variation in alternative hypothesis of evolutionary time for the same organisms, and can foster exploration of the effect of alternative evolutionary timing hypotheses on the results of downstream analyses, providing a framework for a more informed interpretation of evolutionary results.
}

@@

@@

\hypertarget{description}{%
\section{Description}\label{description}}

DateLife's core software application consists of the R package \texttt{datelife}. Its current stable version -- v0.6.6, is available from The Comprehensive R Archive Network (CRAN) repository (Sanchez-Reyes et al., 2022), and relies on functionalities from various biological R packages:
ape (Paradis, Claude, \& Strimmer, 2004),
bold (Chamberlain, 2018),
geiger (Pennell et al., 2014),
msa (Bodenhofer, Bonatesta, Horejš-Kainrath, \& Hochreiter, 2015),
paleotree (Bapst, 2012),
phyloch (Heibl, 2008),
phylocomr (Ooms \& Chamberlain, 2018),
phytools (Revell, 2012),
rotl (Michonneau, Brown, \& Winter, 2016), and
taxize (Chamberlain, 2018; Chamberlain \& Szöcs, 2013).
Figure \ref{fig:workflow} provides a graphical summary of the three main steps of the DateLife workflow: creating a search query, searching a database, and summarizing results from the search.

@@

DateLife processes input scientific names using a Taxonomic Name Resolution Service (TNRS), which increases the probability of correctly finding the queried taxon names in the chronogram database. TNRS detects, corrects and standardizes name misspellings and typos, variant spellings and authorities, and nomenclatural synonyms to a single taxonomic standard (Boyle et al., 2013). TNRS also allows to correctly choose between homonyms, by considering other taxa provided as input to infer the taxonomic context of the homonym. DateLife implements TNRS using the Open Tree of Life (OpenTree) unified Taxonomy (OTT, Open Tree Of Life et al., 2016; Rees \& Cranston, 2017) as standard, storing taxonomic identification numbers (OTT ids) for further processing and analysis. Other taxonomies currently supported by DateLife are the National Center of Biotechnology Information (NCBI) taxonomic database (Schoch et al., 2020), the Global Biodiversity Information Facility (GBIF) taxonomic backbone (GBIF Secretariat, 2022), and the Interim Register of Marine and Nonmarine Genera (IRMNG) database (Rees et al., 2017).

@@

A unique feature of the Phylesystem is that any user can add new published, state-of-the-art chronograms any time, through their curator application (\url{https://tree.opentreeoflife.org/curator}). As chronograms are added to Phylesystem, they can be incorporated into the chronogram database of the \texttt{datelife} package. \texttt{datelife}'s chronogram database is currently manually updated as new chronogram data is added to Phylesystem. The updated database is assigned a new version number, followed by a package release on CRAN.
Users can also implement functions from the \texttt{datelife} R package to trigger an update of the local chronogram database, to incorporate any new chronograms to the user's DateLife analysis before an official database update is released on CRAN.

@@

BOLD trees generated by DateLife, and any tree with branch lengths provided by a user can be analysed with phylogenetic dating methods supported in DateLife that incorporate branch length information from the input topology in combination with the secondary calibrations:
PATHd8, a non-clock, rate-smoothing method to date trees (Britton et al., 2007);
treePL (Smith \& O'Meara, 2012), a semi-parametric, rate-smoothing, penalized likelihood dating method (Sanderson, 2002);
and MRBayes (Huelsenbeck \& Ronquist, 2001; Ronquist \& Huelsenbeck, 2003), a Bayesian inference program implementing Markov chain Monte Carlo (MCMC) methods to estimate a posterior distribution of model parameters.

@@

\texttt{datelife}'s R package code speed was tested on an Apple iMac
with one 3.4 GHz Intel Core i5 processor.
We registered variation in computing time of query processing and search through the database relative to number of queried taxon names.
Query processing time increases roughly linearly with number of input taxon names, and
increases considerably if Taxonomic Name Resolution Service (TNRS) is activated.
Up to ten thousand names can be processed and searched in less than 30 minutes with the most time consuming settings.
Once names have been processed as described in methods, a name search through the chronogram database can be performed in less than a minute, even with a very large number of taxon names (Fig. \ref{fig:runtime_main}).

@@

We illustrate the DateLife workflow using a family within the Passeriform birds encompassing the true finches, Fringillidae, as case study. On a small example, we analysed 6 bird species, and results from each step of the workflow are shown in Fig. \ref{fig:figure2}. As a second example, we analysed 289 bird species in the family Fringillidae that are included in the NCBI taxonomy. The resulting summary chronogram is shown in Fig. \ref{fig:fringillidages}, and results from previous steps of the workflow are available as Supplementary Figures.

@@

To obtain ages for all species within the family of true finches, Fringillidae, we ran a DateLife query using the \enquote{get species from taxon} flag,
which gets all recognized species names within a named group from a taxonomy of choice.
Following the NCBI taxonomy, our DateLife query has 289 Fringillidae species names.
This taxon-constrained approach implies that the full DateLife analysis will be performed using a tree topology and ages available for species names from a given taxonomic group, which do not necessarily correspond to a monophyletic group. Users can change this behaviour by providing all species names corresponding to a monophyletic group as input for a DateLife search, or a monophyletic tree to construct a DateLife summary.

@@

Next, we used the processed species names in our DateLife query to identify chronograms with at least two Fringillidae species as tip taxa.
The DateLife search identified 19 chronograms matching this criteria, published in 13 different studies (Barker et al., 2013, 2015; Burns et al., 2014; Claramunt \& Cracraft, 2015; Gibb et al., 2015; Hedges et al., 2015; Hooper \& Price, 2017; Jetz et al., 2012; Kimball et al., 2019; Oliveros et al., 2019; Price et al., 2014; Roquet, Lavergne, \& Thuiller, 2014; Uyeda, Pennell, Miller, Maia, \& McClain, 2017).
Once identified, DateLife pruned these matching chronograms to remove tips that do not belong to the queried taxon names, and transformed these pruned chronograms to pairwise distance matrices, revealing 1, 206 different age data points available for species within the Fringillidae (Supplementray Table S1).

@@

We performed a cross validation analysis of the DateLife workflow using the Fringillidae chronograms.
We used the individual tree topologies from each of the 19 source chronograms from 13 studies as inputs, treating their node ages as unknown.
We then estimated dates for these topologies using the node ages from the chronograms from the other studies as calibrations and smoothing using BLADJ.
We found that node ages from original study, and ages estimated using all other age data available are correlated (Fig. \ref{fig:cvXY}).
For five studies, Datelife tended to underestimate ages for topologically deeper nodes (those with many descendant taxa, aka \enquote{closer to the root}) relative to the original estimate, and overestimate ages for nodes closer to the tips.
Accordingly, root ages are generally older in the original study than estimated using cross-validated ages (Supplementary Fig. S1).

@@

Conflict in estimated ages among alternative studies is common in the literature. See, for example, the robust ongoing debate about crown group age of angiosperms (Barba-Montoya, Reis, Schneider, Donoghue, \& Yang, 2018; Magallón, Gómez-Acevedo, Sánchez-Reyes, \& Hernández-Hernández, 2015; Ramshaw et al., 1972; Sanderson \& Doyle, 2001; Sauquet, Ramírez-Barahona, \& Magallón, 2021).
Source chronograms available for the same organisms have potentially been estimated implementing calibrations very differently.
For example, the chronograms from Burns et al. (2014) were inferred using molecular substitution rate estimates across birds (Weir \& Schluter, 2008), and have much older age estimates for the same nodes than chronograms that were inferred using fossils as calibrations (Figs. \ref{fig:fringillidages}, \ref{fig:cvXY}; Supplementary Figs. S1, S5).

@@

Taxonomy based politomy resolvers also introduce topological differences.
The study of macroevolutionary processes largely depends on an understanding of the timing of species diversification events, and different phylogenetic and chronogram hypothesis can provide very different overviews of the macroeovlutionary history of a biological group.
For example, alternative topologies in chronograms from the same biological group can infer very different species diversification patterns (Rabosky, 2015; Title \& Rabosky, 2016).
Similarly there are worries that patterns of morphological evolution cannot be accurately inferred with these types of phylogenies, as any patterns would be erased by randomization (Rabosky, 2015). We note that the same applies for geographical and morphological dependent diversification analysis. Hence, we suggest that these types of randomly resolved phylogenies can be used as null or neutral models, representing the case of diversification independent of traits and geographical scenario.

@@

The DateLife software is free and open source. It can be used online through its R shiny web application hosted at \url{http://www.datelife.org}, and locally through the \texttt{datelife} R package, downloadable from Zenodo (\url{https://doi.org/10.5281/zenodo.593938}) and the CRAN repository (Sanchez-Reyes et al., 2022).
DateLife's web application is maintained using RStudio's shiny server and the shiny package open infrastructure, as well as Docker and OpenTree's infrastructure (dates.opentreeoflife.org/datelife).
\texttt{datelife}'s R package stable version is available for installation from CRAN repository using the command \texttt{install.packages(pkgs\ =\ "datelife")} from within R.
Development versions are available from DateLife's GitHub repository (\url{https://github.com/phylotastic/datelife}) and can be installed using the command \texttt{devtools::install\_github("phylotastic/datelife")}.

@@

The DateLife project was born as a prototype tool aiming to provide the services describe in this paper, and was initially developed over a series of hackathons at the National Evolutionary Synthesis Center, NC, USA (Stoltzfus et al., 2013).
We thank colleagues from the O'Meara Lab at the University
of Tennesse Knoxville for suggestions, discussions and software testing.
The late National Evolutionary Synthesis Center (NESCent), which sponsored hackathons
that led to initial work on this project. The team that assembled DateLife's first proof of concept: Tracy Heath, Jonathan Eastman, Peter Midford, Joseph Brown, Matt Pennell, Mike Alfaro, and Luke Harmon.
The Open Tree of Life project that provides the open, metadata rich repository of
trees used to construct DateLife's chronogram database.
The many scientists who publish their chronograms in an open, reusable form, and the scientists who curate them for deposition in the Open Tree of Life repository.
The NSF for funding nearly all the above, in addition to the ABI grant that funded this project itself.

>>>>

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[english,man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={DateLife: leveraging databases and analytical tools to reveal the dated Tree of Life},
            pdfauthor={Luna L. Sánchez Reyes1,2, Emily Jane McTavish1, \& Brian O'Meara2},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

@@

@@

\begin{center}
Abstract
\end{center}

Chronograms --phylogenies with branch lengths proportional to time-- represent key data on evolutionary time frame for the study of natural processes in many areas of biological research.
Chronograms also provide valuable information that can be used for education, science communication, and policy decisions.
Yet, achieving a high-quality reconstruction of a chronogram is a difficult and resource-consuming task.
Here we present DateLife, a service implemented as an R package and an R Shiny web application available at www.datelife.org, that provides services for efficient and easy discovery, summary, reuse, and reanalysis of node age data mined from a curated database of expert, peer-reviewed, and openly available chronograms.
The main DateLife workflow starts with one or more scientific taxon names provided by a user. Names are processed and standardized to a unified taxonomy. Then, DateLife runs a name match across its local chronogram database that is curated from Open Tree of Life's phylogenetic repository, and extracts all chronograms that contain at least two queried taxon names, along with their metadata.
Finally, node ages from matching chronograms are mapped upon a chosen tree topology using the congruification algorithm. Congruified node ages are used as secondary calibrations to date the chosen topology, with or without initial branch lengths, using different phylogenetic dating methods such as BLADJ, treePL, PATHd8 and MrBayes.
We performed a cross-validation test to compare node ages resulting from a DateLife analysis (i.e, phylogenetic dating using secondary calibrations) to those from the original chronograms (i.e, obtained with primary calibrations), and found that DateLife's node age estimates appear generally correlated to those from the original chronograms, with the largest variation in ages occurring around topologically deeper nodes.
Software such as DateLife can help increase awareness of the existing variation in alternative hypothesis of evolutionary time frame for the same organisms, and can support exploration of the effect of alternative chronogram hypotheses on downstream analyses, providing a framework for a more informed interpretation of evolutionary results.

\emph{Keywords}: Tree; Phylogeny; Scaling; Dating; Ages; Divergence times; Open Science; Congruification; Supertree; Calibrations; Secondary calibrations.

Word count: 5956

@@

\begin{center}
\textsc{Description}
\end{center}

DateLife's core software applications are implemented in the R package \texttt{datelife}. Its current stable version -- v0.6.6, is available from The Comprehensive R Archive Network (CRAN) repository (Sanchez-Reyes et al., 2022), and relies on functionalities from other biological R packages:
ape (Paradis, Claude, \& Strimmer, 2004),
bold (Chamberlain, 2018),
geiger (Pennell et al., 2014),
msa (Bodenhofer, Bonatesta, Horejš-Kainrath, \& Hochreiter, 2015),
paleotree (Bapst, 2012),
phyloch (Heibl, 2008),
phylocomr (Ooms \& Chamberlain, 2018),
phytools (Revell, 2012),
rotl (Michonneau, Brown, \& Winter, 2016), and
taxize (Chamberlain, 2018; Chamberlain \& Szöcs, 2013).
Figure \ref{fig:workflow} provides a graphical summary of the three main steps of the DateLife workflow: creating a search query, searching a database, and summarizing results from the search.

@@

DateLife processes input scientific names using a Taxonomic Name Resolution Service (TNRS), which increases the probability of correctly finding the queried taxon names in the chronogram database. TNRS detects, corrects and standardizes name misspellings and typos, variant spellings and authorities, and nomenclatural synonyms to a single taxonomic standard (Boyle et al., 2013). TNRS also allows to correctly choose between homonyms, by considering other taxa provided as input to infer the taxonomic context of the homonym. DateLife implements TNRS using the Open Tree of Life (OpenTree) unified Taxonomy (OTT, Open Tree Of Life et al., 2016; Rees \& Cranston, 2017) as standard, storing taxonomic identification numbers (OTT ids) for further processing and analysis. Other taxonomies currently supported by DateLife are the National Center of Biotechnology Information (NCBI) taxonomic database (Schoch et al., 2020), the Global Biodiversity Information Facility (GBIF) taxonomic backbone (GBIF Secretariat, 2022), and the Interim Register of Marine and Non-marine Genera (IRMNG) database (Rees et al., 2017).

@@

A unique feature of the Phylesystem is that any user can add new published, state-of-the-art chronograms any time, through their curator application (\url{https://tree.opentreeoflife.org/curator}). As chronograms are added to Phylesystem, they can be incorporated into the chronogram database of the \texttt{datelife} R package, which is currently manually updated as new chronogram data is added to Phylesystem. The updated database is assigned a new version number, followed by a package release on CRAN.
Users can directly implement \texttt{datelife} functions to trigger an update of the local chronogram database, to incorporate any new chronograms to their DateLife analysis before an official database update is released on CRAN.

@@

Phylogenetic dating methods supported in DateLife that incorporate branch length information from the input topology in combination with the secondary calibrations include:
PATHd8, a non-clock, rate-smoothing method to date trees (Britton et al., 2007);
treePL (Smith \& O'Meara, 2012), a semi-parametric, rate-smoothing, penalized likelihood dating method (Sanderson, 2002);
and MrBayes (Huelsenbeck \& Ronquist, 2001; Ronquist \& Huelsenbeck, 2003), a Bayesian inference program implementing Markov chain Monte Carlo (MCMC) methods to estimate a posterior distribution of model parameters.

@@

R package \texttt{datelife} code speed was tested on an Apple iMac
with one 3.4 GHz Intel Core i5 processor.
We registered variation in computing time of query processing and search through the database relative to number of queried taxon names.
Query processing time increases roughly linearly with number of input taxon names, and
increases considerably if Taxonomic Name Resolution Service (TNRS) is activated.
Up to ten thousand names can be processed and searched in less than 30 minutes with the most time consuming settings.
Once names have been processed as described in methods, a name search through the chronogram database can be performed in less than a minute, even with a very large number of taxon names (Fig. \ref{fig:runtime_main}).

@@

We illustrate the DateLife workflow using a family within the passeriform birds encompassing the true finches, Fringillidae, as case study. On a small example, we analysed 6 bird species, and results from each step of the workflow are shown in Fig. \ref{fig:figure2}. As a second example, we analysed 289 bird species in the family Fringillidae that are included in the NCBI taxonomy. The resulting summary chronogram is shown in Fig. \ref{fig:fringillidages}, and results from previous steps of the workflow are available as Supplementary Figures.

@@

To obtain ages for all species within the family of true finches, Fringillidae, we ran a DateLife query using the \enquote{get species from taxon} flag,
which gets all recognized species names within a named group from a taxonomy of choice.
Following the NCBI taxonomy, our DateLife query has 289 Fringillidae species names.
This taxon-constrained approach implies that the full DateLife analysis will be performed using a tree topology and ages available for species names from a given taxonomic group, which do not necessarily correspond to a monophyletic group. Users can change this behavior by providing all species names corresponding to a monophyletic group as input for a DateLife search, or a monophyletic tree to construct a DateLife summary.

@@

Next, we used the processed species names in our DateLife query to identify chronograms with at least two Fringillidae species as tip taxa.
The DateLife search identified 19 chronograms matching this criteria, published in 13 different studies (Barker et al., 2013, 2015; Burns et al., 2014; Claramunt \& Cracraft, 2015; Gibb et al., 2015; Hedges et al., 2015; Hooper \& Price, 2017; Jetz et al., 2012; Kimball et al., 2019; Oliveros et al., 2019; Price et al., 2014; Roquet, Lavergne, \& Thuiller, 2014; Uyeda, Pennell, Miller, Maia, \& McClain, 2017).
Once identified, DateLife pruned these matching chronograms to remove tips that do not belong to the queried taxon names, and transformed these pruned chronograms to pairwise distance matrices, revealing 1, 206 different age data points available for species within the Fringillidae (Supplementary Table S1).

@@

We performed a cross validation analysis of the DateLife workflow using the Fringillidae chronograms.
We used the individual tree topologies from each of the 19 source chronograms from 13 studies as inputs, treating their node ages as unknown.
We then estimated dates for these topologies using the node ages from the chronograms from the other studies as calibrations and smoothing using BLADJ.
We found that node ages from original study, and ages estimated using all other age data available are correlated (Fig. \ref{fig:cvXY}).
For five studies, DateLife tended to underestimate ages for topologically deeper nodes (those with many descendant taxa, aka \enquote{closer to the root}) relative to the original estimate, and overestimate ages for nodes closer to the tips.
Accordingly, root ages are generally older in the original study than estimated using cross-validated ages (Supplementary Fig. S1).

@@

Conflict in estimated ages among alternative studies is common in the literature. See, for example, the robust ongoing debate about crown group age of angiosperms (Barba-Montoya, Reis, Schneider, Donoghue, \& Yang, 2018; Magallón, Gómez-Acevedo, Sánchez-Reyes, \& Hernández-Hernández, 2015; Ramshaw et al., 1972; Sanderson \& Doyle, 2001; Sauquet, Ramírez-Barahona, \& Magallón, 2021).
Source chronograms available for the same organisms have potentially been estimated implementing calibrations very differently.
In the Fringillidae example shown here, the chronograms from Burns et al. (2014) were inferred using molecular substitution rate estimates across birds (Weir \& Schluter, 2008), and have much older age estimates for the same nodes than chronograms that were inferred using fossils as calibrations (Figs. \ref{fig:fringillidages}, \ref{fig:cvXY}; Supplementary Figs. S1, S5).

@@

Taxonomy based politomy resolvers also introduce topological differences.
The study of macroevolutionary processes largely depends on an understanding of the timing of species diversification events, and different phylogenetic and chronogram hypothesis can provide very different overviews of the macroevolutionary history of a biological group.
For example, alternative topologies in chronograms from the same biological group can infer very different species diversification patterns (Rabosky, 2015; Title \& Rabosky, 2016).
Similarly there are worries that patterns of morphological evolution cannot be accurately inferred with these types of phylogenies, as any patterns would be erased by randomization (Rabosky, 2015). We note that the same applies for geographical and morphological dependent diversification analysis. Hence, we suggest that these types of randomly resolved phylogenies can be used as null or neutral models, representing the case of diversification independent of traits and geographical scenario.

@@

The DateLife software is free and open source. It can be used online through its R shiny web application hosted at \url{http://www.datelife.org}, and locally through the \texttt{datelife} R package, downloadable from Zenodo (\url{https://doi.org/10.5281/zenodo.593938}) and the CRAN repository (Sanchez-Reyes et al., 2022).
DateLife's web application is maintained using RStudio's shiny server and the shiny package open infrastructure, as well as Docker and OpenTree's infrastructure (dates.opentreeoflife.org/datelife).
\texttt{datelife}'s stable version is available for installation from CRAN repository using the command \texttt{install.packages(pkgs\ =\ "datelife")} from within R.
Development versions are available from DateLife's GitHub repository (\url{https://github.com/phylotastic/datelife}) and can be installed using the command \texttt{devtools::install\_github("phylotastic/datelife")}.

@@

The DateLife project was born as a prototype tool aiming to provide the services describe in this paper, and was initially developed over a series of hackathons at the National Evolutionary Synthesis Center, NC, USA (Stoltzfus et al., 2013).
We thank colleagues from the O'Meara Lab at the University
of Tennessee Knoxville for suggestions, discussions and software testing.
The late National Evolutionary Synthesis Center (NESCent), which sponsored hackathons
that led to initial work on this project. The team that assembled DateLife's first proof of concept: Tracy Heath, Jonathan Eastman, Peter Midford, Joseph Brown, Matt Pennell, Mike Alfaro, and Luke Harmon.
The Open Tree of Life project that provides the open, metadata rich repository of
trees used to construct DateLife's chronogram database.
The many scientists who publish their chronograms in an open, reusable form, and the scientists who curate them for deposition in the Open Tree of Life repository.
The NSF for funding nearly all the above, in addition to the ABI grant that funded this project itself.

### 2023-02-21 review.Rmd
provide a better abstract

@@

Thanks for pointing that out. We added that topic and a couple of references to the intro, L48-49.

>>>>

Thanks for your comment, the abstract was indeed slightly forsaken on the previous version of the ms. We rewrote most of the abstract and checked for spelling. We hope you find it a good improvement from the original.

@@

Thanks for pointing that out. We added that topic to the intro along with a couple of references, L48-49.
